---
title: 'Tutorial: Meta-Analysis of Proportion in R'
author: "Pius S. Ekong"
date: "July 9, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset

Dataset was compiled through a systematic review process

## Install packages required for the analysis
```{r package}
# install.packages("tidyverse")
# install.packages("meta")
# install.packages("metafor")
```

## Load library
```{r library}
library(tidyverse)
library(meta)
library(metafor)
```

## Read the dataset into R
```{r data}
hidedata <- read.csv("Hide_Data.csv", sep="," , header=T, fill=T, na.string=" ")
head(hidedata,10)
```

## Data cleaning
Look at data loaded as "hidedata"
```{r hide}
glimpse(hidedata)
```

Replace the variable names in the dataset
```{r names}
names(hidedata) <- c("ID", "StudyName", "StudySetting", "StudyDesign", "CattleType", "SampleType", "AreaSwab", 
                     "NumSample", "NumPositive", "PropPositive", "Country", "Serotype", "Season")
glimpse(hidedata)
str(hidedata)
```

Let's take a look at the summary of the data
```{r summ}
summary(hidedata)
```

The variable "StudySetting" is suppose to have 2 levels - "Commercial" & "Research Facility"
```{r conv}
convert <- which(hidedata$StudySetting =="commercial")
hidedata$StudySetting[convert] <- "Commercial"
summary(hidedata$StudySetting)
```

Let's get rid of the empty factor level 'commercial'
```{r rid}
hidedata$StudySetting <- droplevels(hidedata$StudySetting)
summary(hidedata$StudySetting)
```

# Data Visualization
Categorical data
```{r catdata}
par(mfrow = c(2,4))
barplot(Serotype <- table(hidedata$StudySetting))
barplot(Serotype <- table(hidedata$StudyDesign))
barplot(Serotype <- table(hidedata$CattleType))
barplot(Serotype <- table(hidedata$SampleType))
barplot(Serotype <- table(hidedata$AreaSwab))
barplot(Serotype <- table(hidedata$Country))
barplot(Serotype <- table(hidedata$Serotype))
barplot(Serotype <- table(hidedata$Season))
```

Numeric data
```{r numdata}
par(mfrow = c(1,3))
hist(hidedata$NumSample)
hist(hidedata$NumPositive)
hist(hidedata$PropPositive)
```

A priori we believe proportion positive differ by CattleType, SampleType and Season.
Take a subset of data for fed beef sampled at the farm 'Hide-farm'
```{r priori}
hidedata_FB_Farm <- subset(hidedata, CattleType=="Fed beef" & SampleType=="Hide-Farm" & Season=="Summer")
dim(hidedata_FB_Farm)
```

## Meta-analysis

```{r Meta-analysis}
# E. coli on hide for fed beef sampled at feedlot 
hidemeta_FB_Farm_S <- metaprop(NumPositive, NumSample, studlab=StudyName, sm="PFT", data=hidedata_FB_Farm, method="Inverse", method.tau="DL")
summary(hidemeta_FB_Farm_S)
```

## Forest plot 
```{r Forest plot}
# Summary effect measure obtained from random effect model only
forest.meta(hidemeta_FB_Farm_S, layout="RevMan5", xlab="Proportion", comb.r=T, comb.f=F, xlim = c(0,1), fontsize=10, digits=3)
```

## Small Study Effects
Assessment for small study effects
```{r Small study}
funnel.meta(hidemeta_FB_Farm_S)
metabias(hidemeta_FB_Farm_S)
```

## Meta-regression: Univariable
```{r Univariable}
metareg(hidemeta_FB_Farm_S, ~ StudyDesign)
metareg(hidemeta_FB_Farm_S, ~ StudySetting)
metareg(hidemeta_FB_Farm_S, ~ AreaSwab)
metareg(hidemeta_FB_Farm_S, ~ Country)
metareg(hidemeta_FB_Farm_S, ~ Serotype)
```

## Meta-regression: Multivariable
```{r Multivariable}
metareg(hidemeta_FB_Farm_S, ~ StudySetting + AreaSwab + Country)
```

# Reference:
Details of the systematic review and meta-analysis process and results was published here.
Ekong P.S., Sanderson M.W., Cernicchiaro N., 2015. Prevalence and concentration of Escherichia coli O157 in different seasons and cattle types processed in North America: A systematic review and meta-analysis of published research. Prev. Vet. Med. 121, 74–85
